# LevenR --- Calculate Levenshtein distance between strings

[![Build Status](https://travis-ci.org/sherrillmix/levenR.svg?branch=master)](https://travis-ci.org/sherrillmix/levenR)

## Introduction

`violinPointR` provides a few functions for simple Levenshtein alignment and distance calculation with multiple threads, ends-free and reduced homopolymer gap costs.

## Installation
To install directly from github, use the [<code>devtools</code>](https://github.com/hadley/devtools) library and run:
```{r, eval=FALSE}
devtools::install_github("sherrillmix/levenR")
```

## Examples

### Generate Levenshtein distance matrix

An example of calculating the Levenshtein distance between several strings to make a distance matrix:
```{r levenMatrix, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
leven(seqs)
```

### Compare many to one

An example of calculating the Levenshtein distance between several strings against a longer reference sequence:
```{r levenRef, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
ref<-'CCAAATACCGACC'
leven(seqs,ref,substring2=TRUE)
```

### Find the best reference

An example of calculating the Levenshtein distance between several strings against two longer reference sequences and determining the best match for each read:
```{r bestRef, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
refs<-c('CCATAATACCGACC','GGAAATACCTA')
dist<-leven(seqs,refs,substring2=TRUE)
apply(dist,1,which.min)
```

### Ignoring indels in homopolymers
An example of calculating the Levenshtein distance between several strings to make a distance matrix while ignoring indels in long homopolymers (an error type common in 454 and IonTorrent sequencing):
```{r levenHomo, tidy=TRUE}
library(levenR)
seqs<-c('AAAAATA','AAATTTTTA','AAAAATTTA')
leven(seqs,homoLimit=3)
```

### Using multiple threads 
An example of calculating the Levenshtein distance between several strings using multiple threads:
```{r levenThreads, tidy=TRUE}
library(levenR)
seqs<-replicate(50,paste(sample(letters,100,TRUE),collapse=''))
system.time(leven(seqs))
system.time(leven(seqs,nThreads=4))
```


### Alignment

An example of aligning strings against a longer reference:
```{r align, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
ref<-'CCAAATACCGACC'
levenAlign(seqs,ref,substring2=TRUE)
```



