# Calculate Levenshtein distance between strings

## Introduction

`violinPointR` provides a few functions for simple Levenshtein alignment and distance calculation with multiple threads, ends-free and reduced homopolymer gap costs.

## Installation

```{r, eval=FALSE}
devtools::install_github("sherrillmix/levenR")
```

## Examples

### Generate Levenshtein distance matrix

An example of calculating the Levenshtein distance between several strings to make a distance matrix:
```{r levenMatrix, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
leven(seqs)
```

### Compare many to one

An example of calculating the Levenshtein distance between several strings against a longer reference sequence:
```{r levenRef, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
ref<-'CCAAATACCGACC'
leven(seqs,ref,substring2=TRUE)
```

### Find the best reference

An example of calculating the Levenshtein distance between several strings against two longer reference sequences and determining the best match for each read:
```{r bestRef, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
refs<-c('CCATAATACCGACC','GGAAATACCTA')
dist<-leven(seqs,refs,substring2=TRUE)
apply(dist,1,which.min)
```

### Alignment

An example of aligning strings against a longer reference:
```{r align, tidy=TRUE}
library(levenR)
seqs<-c('AAATA','AATA','AAAT','ACCTA')
ref<-'CCAAATACCGACC'
levenAlign(seqs,ref,substring2=TRUE)
```



